---
title: "MRS Longitudinal Modelling"
author: "Meaghan Perdue"
date: "12/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(root.dir = '/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/')
library(tidyverse)
library(lme4)
library(lmerTest)
library(ggplot2)
library(ggpubr)
library(EnvStats)
library(psych)
library(viridis)
library(remef)

setwd('/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data')
```

## Organize data

Merge output data from multiple batches, select metabolite columns, add participant info and other data

```{r}
acg_batch1 <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/ACG_LCM_CorrectedTable_12-14-2021-16-15.csv")

acg_batch2 <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/ACG_LCM_CorrectedTable_12-14-2021-16-21.csv")

acg_LCM_corr <- bind_rows(acg_batch1, acg_batch2)
anyDuplicated(acg_LCM_corr$File_ID) #check for duplicated values
acg_LCM_corr<-unique(acg_LCM_corr) #remove duplicate rows
acg_LCM_corr["4", "Cho_GPC_PCh_conc_molar"]<-NA #Cho conc invalid for PS10008_PS14_046_2386 due to spike in spectrum
acg_LCM_corr["4", "Cho_GPC_PCh_conc_molal"]<-NA #Cho conc invalid for PS10008_PS14_046_2386 due to spike in spectrum

write.csv(acg_LCM_corr, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/ACG_LCM_CorrectedTable_12-14-2021_all.csv", row.names = FALSE)

acg_molal <- acg_LCM_corr %>%
  dplyr::select(File_ID, fGM, fWM, fCSF, NAA_conc_molal, Cre_PCr_conc_molal, Cho_GPC_PCh_conc_molal, Ins_conc_molal, Glu_conc_molal, Glu_Gln_conc_molal)

acg_demo <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/acg_analysis_subs.csv")

acg_dat <- left_join(acg_demo, acg_molal, by = "File_ID")
write.csv(acg_dat, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/acg_data_molal_n113.csv", row.names = FALSE)

lag_batch1 <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/LAG_LCM_CorrectedTable_12-13-2021-15-18.csv")

lag_batch2 <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/LAG_LCM_CorrectedTable_12-14-2021-11-41.csv")

lag_batch3 <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/LAG_LCM_CorrectedTable_12-14-2021-12-18.csv")

lag_LCM_corr <- bind_rows(lag_batch1, lag_batch2, lag_batch3)
write.csv(lag_LCM_corr, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/LAG_LCM_CorrectedTable_12-14-2021_all.csv", row.names = FALSE)

lag_molal <- lag_LCM_corr %>%
  dplyr::select(File_ID, fGM, fWM, fCSF, NAA_conc_molal, Cre_PCr_conc_molal, Cho_GPC_PCh_conc_molal, Ins_conc_molal, Glu_conc_molal, Glu_Gln_conc_molal)

lag_demo <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/lag_analysis_subs.csv")

lag_dat <- left_join(lag_demo, lag_molal, by = "File_ID")
lag_dat$subj_id<-as.character(lag_dat$subj_id) #make subj_id character type

write.csv(lag_dat, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/lag_data_molal_n321.csv", row.names = FALSE)

```
##Check data distribution & outliers
Create plots to check data distribution and identify outliers

ACG
```{r}
plot_age <- ggplot(acg_dat, aes(mri_age_y))
plot_age + geom_density()
#age skewed

plot_naa <- ggplot(acg_dat, aes(NAA_conc_molal))
plot_naa + geom_density()
plot_naa + geom_boxplot()
rosnerTest(acg_dat$NAA_conc_molal, k = 4)

plot_cre <- ggplot(acg_dat, aes(Cre_PCr_conc_molal))
plot_cre + geom_density()
plot_cre + geom_boxplot()
rosnerTest(acg_dat$Cre_PCr_conc_molal, k = 5)

plot_cho <- ggplot(acg_dat, aes(Cho_GPC_PCh_conc_molal))
plot_cho + geom_density()
plot_cho + geom_boxplot()
rosnerTest(acg_dat$Cho_GPC_PCh_conc_molal, k = 4)

plot_ins <- ggplot(acg_dat, aes(Ins_conc_molal))
plot_ins + geom_density()
shapiro.test(acg_dat$Ins_conc_molal) #not normal
plot_ins + geom_boxplot()
rosnerTest(acg_dat$Ins_conc_molal, k = 3) #outlier value 1.471244, observation 84 (PS18_013) - spectrum showed issue, exclude from analysis 
acg_dat["84", "Ins_conc_molal"] <- NA #replace outlier value with NA to exclude from analysis

plot_glx <- ggplot(acg_dat, aes(Glu_Gln_conc_molal))
plot_glx + geom_density()
plot_glx + geom_boxplot()
rosnerTest(acg_dat$Glu_Gln_conc_molal, k = 1)

write.csv(acg_dat, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/acg_data_molal_outliers_removed_n113.csv", row.names = FALSE)
```
LAG
```{r}
plot_age <- ggplot(lag_dat, aes(mri_age_y))
plot_age + geom_density()
#age skewed

plot_naa <- ggplot(lag_dat, aes(NAA_conc_molal))
plot_naa + geom_density()
plot_naa + geom_boxplot()
rosnerTest(lag_dat$NAA_conc_molal, k = 10) #outliers val. 19.207, obs. 296 & Val. 10.148, obs. 306 - some issues in spectra, exclude from analysis for NAA
lag_dat["296", "NAA_conc_molal"]<- NA
lag_dat["306", "NAA_conc_molal"]<- NA

plot_cre <- ggplot(lag_dat, aes(Cre_PCr_conc_molal))
plot_cre + geom_density()
plot_cre + geom_boxplot()
rosnerTest(lag_dat$Cre_PCr_conc_molal, k = 10) #outliers val. 14.829, obs. 241 (PS17_047) - structure in resid. & val. 14.09, obs. 178 (PS17_045, spectrum looked fine), excluding only PS17_047 for now
lag_dat["241", "Cre_PCr_conc_molal"]<-NA  #replace outlier value with NA to exclude from analysis

plot_cho <- ggplot(lag_dat, aes(Cho_GPC_PCh_conc_molal))
plot_cho + geom_density()
plot_cho + geom_boxplot()
rosnerTest(lag_dat$Cho_GPC_PCh_conc_molal, k = 8)

plot_ins <- ggplot(lag_dat, aes(Ins_conc_molal))
plot_ins + geom_density()
plot_ins + geom_boxplot()
rosnerTest(lag_dat$Ins_conc_molal, k = 8) #outliers: val. 14.077, obs. 241 (PS17_047) - structure in resid; val. 11.777, obs. 221 (PS16_014) - spectrum fine, keep for now; val. 1.542, obs. 197 (PS18_028) - noisy spectrum, spiky residuals, exclude; val. 2.04, obs. 62 (PS14_111) - spectrum fine, keep for now
lag_dat["241", "Ins_conc_molal"] <- NA #replace outlier value with NA to exclude from analysis
lag_dat["197", "Ins_conc_molal"] <- NA

plot_glx <- ggplot(lag_dat, aes(Glu_Gln_conc_molal))
plot_glx + geom_density()
plot_glx + geom_boxplot()
rosnerTest(lag_dat$Glu_Gln_conc_molal, k = 7)#outliers: val. 10.09, obs. 241 (PS17_047) - structure in resid; val. 31.2899, obs. 296 (PS21_030) - big spike in resid. around 2.1, exclude.
lag_dat["241", "Glu_Gln_conc_molal"] <- NA
lag_dat["296", "Glu_Gln_conc_molal"] <- NA

#exclude PS17_047 from all analyses due to outlier status on multiple metabolites and issues in spectrum (noisy, structure in resid)
lag_dat["241", "NAA_conc_molal"] <- NA
lag_dat["241", "Cre_PCr_conc_molal"] <- NA
lag_dat["241", "Cho_GPC_PCh_conc_molal"] <- NA

write.csv(lag_dat, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/lag_data_molal_outliers_removed_n320.csv", row.names = FALSE)
```

##Descriptive stats
```{r}
#ACG - count individual subject N and N of females
acg_sub_sex <- acg_dat %>%
  dplyr::select(subj_id, female)
acg_unique_subs<-unique(acg_sub_sex)
length(acg_unique_subs$subj_id)
sum(acg_unique_subs$female)

describe(acg_dat)

#LAG - count individual subject N and N of females
lag_sub_sex <- lag_dat %>%
  dplyr::select(subj_id, female)
lag_unique_subs<-unique(lag_sub_sex)
length(lag_unique_subs$subj_id)
sum(lag_unique_subs$female)

describe(lag_dat)

#Calculate total Ns from both datasets combined
all_dat<-bind_rows(acg_dat, lag_dat)
length(unique(all_dat$File_ID))
length(unique(all_dat$subj_id))
all_sub_sex <- all_dat %>%
  dplyr::select(subj_id, female)
all_unique_subs<-unique(all_sub_sex)
sum(all_unique_subs$female)

describe(all_dat)
```


## LME modelling ACG
run mixed-effects models for ACG with participant as random effect, age, sex and age x sex interaction as fixed effects
NAA
run on acg_data_molal_outliers_removed_n113.csv
```{r}
acg_naa_age<- lmer(NAA_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_naa_age)

acg_naa_age_sex<- lmer(NAA_conc_molal ~ mri_age_y + female + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_naa_age_sex)

anova(acg_naa_age,acg_naa_age_sex)

acg_naa_ageXsex<- lmer(NAA_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_naa_ageXsex)

anova(acg_naa_age,acg_naa_ageXsex)

#including sex in model does not significantly improve fit 
```
Cre
```{r}
acg_cre_age<- lmer(Cre_PCr_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_cre_age)

acg_cre_age_sex<- lmer(Cre_PCr_conc_molal ~ mri_age_y + female + (1 | subj_id), data=acg_dat,  REML=FALSE)
summary(acg_cre_age_sex)

anova(acg_cre_age,acg_cre_age_sex)

acg_cre_ageXsex<- lmer(Cre_PCr_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_cre_ageXsex)

anova(acg_cre_age,acg_cre_ageXsex)
```
Cho
```{r}
acg_cho_age<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat,  REML=FALSE)
summary(acg_cho_age)

acg_cho_age_sex<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + female + (1 | subj_id), data=acg_dat,  REML=FALSE)
summary(acg_cho_age_sex)

anova(acg_cho_age,acg_cho_age_sex)

acg_cho_ageXsex<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=acg_dat,  REML=FALSE)
summary(acg_cho_ageXsex)

anova(acg_cho_age,acg_cho_ageXsex)

#including sex in model does not significantly improve fit 
```
Ins
```{r}
acg_ins_age<- lmer(Ins_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_ins_age)

acg_ins_age_sex<- lmer(Ins_conc_molal ~ mri_age_y + female + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_ins_age_sex)

anova(acg_ins_age,acg_ins_age_sex)

acg_ins_ageXsex<- lmer(Ins_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_ins_ageXsex)

anova(acg_ins_age,acg_ins_ageXsex)
```

Glx
```{r}
acg_glx_age<- lmer(Glu_Gln_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_glx_age)

acg_glx_age_sex<- lmer(Glu_Gln_conc_molal ~ mri_age_y + female + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_glx_age_sex)

anova(acg_glx_age,acg_glx_age_sex)

acg_glx_ageXsex<- lmer(Glu_Gln_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=acg_dat, REML=FALSE)
summary(acg_glx_ageXsex)

anova(acg_glx_age,acg_glx_ageXsex)
```

## Plot results ACG
Create scatter plot with regression line based on model predictions
NAA
```{r}
acg_dat <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/acg_data_molal_outliers_removed_n113.csv")

#pull intercept and slope from model for regression line
fixef.acg_naa<-fixef(acg_naa_age)

#plot
plot_acg_naa_age<-ggplot(acg_dat, aes(x=mri_age_y, y=NAA_conc_molal,color=as.factor(female)))
plot_acg_naa_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Anterior Cingulate", x = "Age (years)", y = "tNAA (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.acg_naa[[1]], slope= fixef.acg_naa[[2]])) + 
  theme_classic()
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/acg_naa_age_viridis.png")


```
Cho
```{r}
#pull intercept and slope from model for regression line
fixef.acg_cho<-fixef(acg_cho_age)

#plot
plot_acg_cho_age<-ggplot(acg_dat, aes(x=mri_age_y, y=Cho_GPC_PCh_conc_molal,color=as.factor(female)))
plot_acg_cho_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Anterior Cingulate", x = "Age (years)", y = "tCho (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.acg_cho[[1]], slope= fixef.acg_cho[[2]])) + 
  theme_classic()
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/acg_cho_age_viridis.png")
```
Ins
```{r}
#pull intercept and slope from model for regression line
fixef.acg_ins<-fixef(acg_ins_age)

#plot
plot_acg_ins_age<-ggplot(acg_dat, aes(x=mri_age_y, y=Ins_conc_molal,color=as.factor(female)))
plot_acg_ins_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Anterior Cingulate", x = "Age (years)", y = "Ins (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.acg_ins[[1]], slope= fixef.acg_ins[[2]])) + 
  theme_classic()
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/acg_ins_age_viridis.png")
```


## LME modelling LAG
run mixed-effects models for lag with participant as random effect, age, sex and age x sex interaction as fixed effects
NAA, substantial longitudinal data so we include random slope
run on lag_data_molal_outliers_removed_n320.csv
```{r}
#basic model effect of age on intercept
lag_naa_age<- lmer(NAA_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age)

lag_naa_age_sex<- lmer(NAA_conc_molal ~ mri_age_y + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age_sex)

anova(lag_naa_age,lag_naa_age_sex)

lag_naa_ageXsex<- lmer(NAA_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_ageXsex)

anova(lag_naa_age,lag_naa_ageXsex)

#including sex in model does not significantly improve fit 

#test quadratic effect of age
#create age^2  variable
lag_dat$mri_age_y2 <- (lag_dat$mri_age_y)^2

lag_naa_age2<- lmer(NAA_conc_molal ~ mri_age_y + mri_age_y2 + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age2) #age^2 p=.0447

anova(lag_naa_age, lag_naa_age2) #fit improved p=.044

```
Cre
```{r}
lag_cre_age<- lmer(Cre_PCr_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age)

lag_cre_age_sex<- lmer(Cre_PCr_conc_molal ~ mri_age_y + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age_sex)

anova(lag_cre_age,lag_cre_age_sex)

lag_cre_ageXsex<- lmer(Cre_PCr_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_ageXsex)

anova(lag_cre_age,lag_cre_ageXsex)
```
Cho
```{r}
lag_cho_age<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age)

lag_cho_age_sex<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age_sex)

anova(lag_cho_age,lag_cho_age_sex)

lag_cho_ageXsex<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_ageXsex)

anova(lag_cho_age,lag_cho_ageXsex)

#including sex in model does not significantly improve fit 
```
Ins
```{r}
lag_ins_age<- lmer(Ins_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age)

lag_ins_age_sex<- lmer(Ins_conc_molal ~ mri_age_y + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age_sex)

anova(lag_ins_age,lag_ins_age_sex)

lag_ins_ageXsex<- lmer(Ins_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_ageXsex)

anova(lag_ins_age,lag_ins_ageXsex)
```

Glx
```{r}
lag_glx_age<- lmer(Glu_Gln_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
isSingular(lag_glx_age) # TRUE: model is almost/near singular (parameters are on the boundary of the feasible parameter space: random effects variance = 0)
summary(lag_glx_age)

lag_glx_age_sex<- lmer(Glu_Gln_conc_molal ~ mri_age_y + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age_sex)

anova(lag_glx_age,lag_glx_age_sex)

lag_glx_ageXsex<- lmer(Glu_Gln_conc_molal ~ mri_age_y + female + mri_age_y:female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_ageXsex)

anova(lag_glx_age,lag_glx_ageXsex)
```

## Plot results LAG
Create scatter plot with regression line based on model predictions
NAA
```{r}
lag_dat<-read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/lag_data_molal_outliers_removed_n320.csv")

#pull intercept and slope from model for regression line
fixef.lag_naa<-fixef(lag_naa_age)

#plot
plot_lag_naa_age<-ggplot(lag_dat, aes(x=mri_age_y, y=NAA_conc_molal,color=as.factor(female)))
plot_lag_naa_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "tNAA (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_naa[[1]], slope= fixef.lag_naa[[2]])) + 
  theme_classic()
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_naa_age_viridis.png")
```

Cho
```{r}
#pull intercept and slope from model for regression line
fixef.lag_cho<-fixef(lag_cho_age)

#plot
plot_lag_cho_age<-ggplot(lag_dat, aes(x=mri_age_y, y=Cho_GPC_PCh_conc_molal,color=as.factor(female)))
plot_lag_cho_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "tCho (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_cho[[1]], slope= fixef.lag_cho[[2]])) + 
  theme_classic()
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_cho_age_viridis.png")

```

Glx
```{r}
#pull intercept and slope from model for regression line
fixef.lag_glx<-fixef(lag_glx_age)

#plot
plot_lag_glx_age<-ggplot(lag_dat, aes(x=mri_age_y, y=Glu_Gln_conc_molal,color=as.factor(female)))
plot_lag_glx_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "Glx (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_glx[[1]], slope= fixef.lag_glx[[2]])) + 
  theme_classic()
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_glx_age_viridis.png")

```

##Combine results plots into 1 figure
Simplify plots to legend & region don't repeat in each panel
```{r}
library(patchwork)
p1_acg_naa_age<-plot_acg_naa_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Anterior Cingulate", x = "Age (years)", y = "tNAA (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.acg_naa[[1]], slope= fixef.acg_naa[[2]])) + 
  theme_classic() +
  theme(plot.title= element_text(hjust = 0.5))

p4_lag_naa_age<-plot_lag_naa_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "tNAA (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  theme_classic() +
  #theme(legend.position = c(0.7, 0.95), legend.background = element_rect(fill='transparent')) +
  geom_abline(aes(intercept = fixef.lag_naa[[1]], slope= fixef.lag_naa[[2]])) +
  theme(plot.title= element_text(hjust = 0.5))


p2_acg_cho_age<-plot_acg_cho_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(x = "Age (years)", y = "tCho (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.acg_cho[[1]], slope= fixef.acg_cho[[2]])) + 
  theme_classic() 
  

p5_lag_cho_age<-plot_lag_cho_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(x = "Age (years)", y = "tCho (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_cho[[1]], slope= fixef.lag_cho[[2]])) + 
  theme_classic()

p3_acg_ins_age<-plot_acg_ins_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(x = "Age (years)", y = "Ins (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.acg_ins[[1]], slope= fixef.acg_ins[[2]])) + 
  theme_classic()

p6_lag_glx_age<-plot_lag_glx_age + 
  geom_point() +
  geom_line(aes(group = as.factor(subj_id))) +
  labs(x = "Age (years)", y = "Glx (molal)") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_glx[[1]], slope= fixef.lag_glx[[2]])) + 
  theme_classic()

results_fig<-p1_acg_naa_age + p4_lag_naa_age + p2_acg_cho_age + p5_lag_cho_age + p3_acg_ins_age + p6_lag_glx_age + plot_layout(nrow=3, guides ='collect')

ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/results_fig_grid_viridis.png", results_fig, width = 2000, units = "px")

```

##Correct for multiple comparisons
Use FDR method (Benjamini & Hochberg) to correct for mutliple comparisons (5 metabolites x 2 voxels = 10 models)
```{r}
#list models (voxel/metabolite pair)
models<- c("acg_naa", "acg_cr", "acg_cho", "acg_ins", "acg_glx", "lag_naa", "lag_cr", "lag_cho", "lag_ins", "lag_glx")

#list p-values for each model, matching order of models
p<-c(.0217, .307, .0258, .0494, .23, .00093, .115, .00229, .636, 0)

fdr_table<-data.frame(models, p)

fdr_table$p_fdr<-p.adjust(p, method = "fdr", n= length(p))
fdr_table

#run FDR without LAG-Glx in case it's invalid:
p2<-c(.0217, .307, .0258, .0494, .23, .00093, .115, .00229, .636)
p.adjust(p2, method = "fdr", n= length(p2))

```


##Additional models
Next steps following OHBM abstract submission (Dec. 2021): test random slopes (LAG only) 

```{r}
#read in data
lag_dat<-read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/lag_data_molal_outliers_removed_n320.csv")

#basic age model NAA
lag_naa_age<- lmer(NAA_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age)

#test random slope NAA
lag_naa_age_randslope <- lmer(NAA_conc_molal ~ mri_age_y + (1 + mri_age_y | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age_randslope)
#model failed to converge 

#basic age model Cre
lag_cre_age<- lmer(Cre_PCr_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age)

#test random slope Cre
lag_cre_age_randslope<- lmer(Cre_PCr_conc_molal ~ mri_age_y + (1 + mri_age_y | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age_randslope)
#model failed to converge

#basic age model Cho
lag_cho_age<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age)

#test random slope Cho
lag_cho_age_randslope<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + (1 + mri_age_y | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age_randslope)

#compare models
anova(lag_cho_age, lag_cho_age_randslope) #fit did not improve with inclusion of random slopes

#basic age model Ins
lag_ins_age<- lmer(Ins_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age)

#test random slope Ins
lag_ins_age_randslope<- lmer(Ins_conc_molal ~ mri_age_y + (1 + mri_age_y | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age_randslope)
#boundary fit is singular

anova(lag_ins_age, lag_ins_age_randslope) #fit did not improve with inclusion of random slopes

lag_glx_age<- lmer(Glu_Gln_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age)
#boundary fit is singular

lag_glx_age_randslope<- lmer(Glu_Gln_conc_molal ~ mri_age_y + (1 + mri_age_y | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age_randslope)
#boundary fit is singular
anova(lag_glx_age, lag_glx_age_randslope) #fit did not improve with inclusion of random slopes

```

##Tissue fraction analysis
Calculate tissue fraction GM/(GM+WM) for each subject/voxel
Add tissue fraction as fixed effect in lmer models
ACG
```{r}
library(ggplot2)
library(EnvStats)
library(lme4)
library(lmerTest)
#ACG
#read in data
acg_dat <- read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/acg_data_molal_outliers_removed_n113.csv")

#create tissue fraction variable
acg_dat$TissueFraction_GMWM <- acg_dat$fGM/(acg_dat$fGM + acg_dat$fWM)

#plot tissue fraction to check for issues
plot_ACG_tissuefraction <- ggplot(acg_dat, aes(TissueFraction_GMWM))
plot_ACG_tissuefraction+ geom_density() #big skew
plot_ACG_tissuefraction + geom_boxplot() #4 potential outliers
rosnerTest(acg_dat$TissueFraction_GMWM, k = 4)
#4 outliers: Obs. Num 37 (PS14_090), 24 (PS14_050), 93 (PS18_1601), 5 (PS14_009) - double check cogregistration
#PS14_090 is shifted to the left and includes substantial WM - EXCLUDE
#PS14_050 is shifted to the left and T1 quality is poor - EXCLUDE
#PS18_1601 head tilted so voxel captures a bit of WM in the left hemi., a little anterior; doesn't seem terrible
#PS14_009 looks like chin was tipped down toward chest a bit, corner of voxel captures some CC white matter but seems ok

#save database with tissue fraction
write.csv(acg_dat, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/acg_data_molal_outliers_removed_n113_tissuefraction.csv", row.names = FALSE)

#exclude scans PS14_090 and PS14_050 due to outlier status coregistration issues
acg_dat_tf<-read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/acg_data_molal_outliers_removed_n113_tissuefraction.csv")
acg_dat_tf["37", "TissueFraction_GMWM"] <- NA
acg_dat_tf["24", "TissueFraction_GMWM"] <- NA

acg_dat_tf <- subset(acg_dat_tf, study_code != c("PS14_090", "PS14_050"))


#check correlation between tissue fraction and age
acg_TisFrac_age <- lmer(TissueFraction_GMWM ~ mri_age_y + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_TisFrac_age)
#not sig.

#Add tissue fraction as fixed effect in ACG models
#NAA
#basic age model
acg_naa_age <- lmer(NAA_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_naa_age)

#add tissue fraction
acg_naa_age_tf <- lmer(NAA_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_naa_age_tf) #no sig. effect of tissue fraction

#compare models
anova(acg_naa_age, acg_naa_age_tf) #model fit not sig. different

#Cre
#basic age model
acg_cre_age <- lmer(Cre_PCr_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_cre_age)

#add tissue fraction
acg_cre_age_tf <- lmer(Cre_PCr_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_cre_age_tf)

#Cho
#basic age model
acg_cho_age <- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_cho_age)

#add tissue fraction
acg_cho_age_tf <- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_cho_age_tf)

#compare models
anova(acg_cho_age, acg_cho_age_tf) #fit not significantly different

#Ins
#basic age model
acg_ins_age <- lmer(Ins_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_ins_age)

#add tissue fraction
acg_ins_age_tf <- lmer(Ins_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_ins_age_tf)

#Glx
#basic age model
acg_glx_age <- lmer(Glu_Gln_conc_molal ~ mri_age_y + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_glx_age)

#add tissue fraction
acg_glx_age_tf <- lmer(Glu_Gln_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=acg_dat_tf, REML=FALSE)
summary(acg_glx_age_tf)



```

LAG Tissue Fraction models
```{r}
#LAG
#read in data
lag_dat<-read.csv("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/lag_data_molal_outliers_removed_n320.csv")

#create tissue fraction variable
lag_dat$TissueFraction_GMWM <- lag_dat$fGM/(lag_dat$fGM + lag_dat$fWM)

#plot tissue fraction to check for issues
plot_LAG_tissuefraction <- ggplot(lag_dat, aes(TissueFraction_GMWM))
plot_LAG_tissuefraction+ geom_density() #normal
plot_LAG_tissuefraction + geom_boxplot() #6 potential outliers
rosnerTest(lag_dat$TissueFraction_GMWM, k = 6) #no significant outliers

write.csv(lag_dat, "/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/data/lag_data_molal_outliers_removed_n320_tissuefraction.csv")

#check correlation between tissue fraction and age
lag_TisFrac_age <- lmer(TissueFraction_GMWM ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_TisFrac_age)
#tissue fraction is significantly negatively associated with age

#in a simple correlation test
cor.test(lag_dat$mri_age_y, lag_dat$TissueFraction_GMWM) # also significant: r = -.37


#Add tissue fraction as a fixed effect in LAG models
#basic age model NAA
lag_naa_age<- lmer(NAA_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)

#add tissue fraction
lag_naa_age_TisFrac <- lmer(NAA_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age_TisFrac)

#compare model fit
anova(lag_naa_age, lag_naa_age_TisFrac) #fit significantly improved

#add tissue fraction interaction with age
lag_naa_age_TisFracXage <- lmer(NAA_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age_TisFracXage)
#interaction term for agexTissueFraction not sig.

#compare model fit
anova(lag_naa_age_TisFrac, lag_naa_age_TisFracXage) #not sig., stick with age + TissueFraction model

#add sex to tissue fraction model
lag_naa_age_TisFrac_sex <- lmer(NAA_conc_molal ~ mri_age_y + TissueFraction_GMWM + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age_TisFrac_sex) #sex not significant

#compare model fit
anova(lag_naa_age_TisFrac, lag_naa_age_TisFrac_sex) #not sig., stick with age + TissueFraction model

#Cre
#basic age model
lag_cre_age<- lmer(Cre_PCr_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age)

#add tissue fraction
lag_cre_age_TisFrac<- lmer(Cre_PCr_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age_TisFrac)

#compare model fit
anova(lag_cre_age, lag_cre_age_TisFrac) #sig.

#add interaction
lag_cre_age_TisFrac_int<- lmer(Cre_PCr_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age_TisFrac_int)

#compare model fit
anova(lag_cre_age_TisFrac, lag_cre_age_TisFrac_int) #not sig.

#add sex to tissue fraction model

lag_cre_age_TisFrac_sex<- lmer(Cre_PCr_conc_molal ~ mri_age_y + TissueFraction_GMWM + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age_TisFrac_sex)

#compare model fit
anova(lag_cre_age_TisFrac, lag_cre_age_TisFrac_sex) #not sig.

#Cho
#basic age model
lag_cho_age<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age)

#add tissue fraction
lag_cho_age_TisFrac<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age_TisFrac)

#compare model fit
anova(lag_cho_age, lag_cho_age_TisFrac) #sig

#add interaction
lag_cho_age_TisFrac_int<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age_TisFrac_int) #interaction is sig. (p=.046)

#compare model fit
anova(lag_cho_age_TisFrac, lag_cho_age_TisFrac_int) #sig. (p<.046)

#add sex to model
lag_cho_age_TisFrac_int_sex<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y * TissueFraction_GMWM + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age_TisFrac_int_sex) #not sig.

#compare model fit
anova(lag_cho_age_TisFrac_int, lag_cho_age_TisFrac_int_sex) #not sig.

#Ins
#basic age model
lag_ins_age<- lmer(Ins_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age)

#add tissue fraction
lag_ins_age_TisFrac<- lmer(Ins_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age_TisFrac) #tissue fraction sig.; age not sig.

#compare model fit
anova(lag_ins_age, lag_ins_age_TisFrac) #sig.

#add interaction

lag_ins_age_TisFrac_int<- lmer(Ins_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age_TisFrac_int) #tissue fraction, age & interaction all sig.

#compare model fit
anova(lag_ins_age_TisFrac, lag_ins_age_TisFrac_int) #sig.

#add sex
lag_ins_age_TisFrac_int_sex<- lmer(Ins_conc_molal ~ mri_age_y * TissueFraction_GMWM + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age_TisFrac_int_sex) #tissue fraction, age & interaction all sig.

#compare model fit
anova(lag_ins_age_TisFrac_int, lag_ins_age_TisFrac_int_sex) #not sig.

#Glx
#basic age model
lag_glx_age<- lmer(Glu_Gln_conc_molal ~ mri_age_y + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age)

#add tissue fraction
lag_glx_age_TisFrac<- lmer(Glu_Gln_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age_TisFrac)

#compare models
anova(lag_glx_age, lag_glx_age_TisFrac) #sig

#add interaction
lag_glx_age_TisFrac_int<- lmer(Glu_Gln_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age_TisFrac_int)

#compare models
anova(lag_glx_age_TisFrac, lag_glx_age_TisFrac_int) #not sig

#add sex
lag_glx_age_TisFrac_sex<- lmer(Glu_Gln_conc_molal ~ mri_age_y + TissueFraction_GMWM + female + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age_TisFrac_sex)

#compare models
anova(lag_glx_age_TisFrac, lag_glx_age_TisFrac_sex) #not sig

#check models for multicollinearity
library(performance)

check_collinearity(lag_naa_age_TisFrac) #Low
check_collinearity(lag_cre_age_TisFrac) #Low
check_collinearity(lag_cho_age_TisFrac_int) #interaction term causes inflated VIFs - check model without interaction
check_collinearity(lag_cho_age_TisFrac) #Low
check_collinearity(lag_ins_age_TisFrac) #Low
check_collinearity(lag_glx_age_TisFrac) #Low


```

##Unpack interactions
Use sim_slopes function from interactions package to conduct simple slopes analysis on continuous x continuous interactions
```{r}
library(interactions)

#LAG Cho: age x tissue fraction interaction
lag_cho_age_TisFrac_int<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)

#run simple slopes analysis with Johnson-Neyman plot
sim_slopes(lag_cho_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM, jnplot = TRUE)

#save Johnson-Neyman plot
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_cho_age_tf_int_JNplot.png")

#create interaction plot (make it pretty later)
#basic plot
Cho_age_tf_int_plot <- interact_plot(lag_cho_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM)
Cho_age_tf_int_plot

#linearity check
Cho_age_tf_int_plot_lin <- interact_plot(lag_cho_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM, modx.values="terciles", plot.points=TRUE, linearity.check=TRUE)
Cho_age_tf_int_plot_lin
#seems to be some non-linear effects, try adding quadratic term

#LAG Ins: age x tissue fraction interaction
lag_ins_age_TisFrac_int<- lmer(Ins_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)

#run simple slopes analysis with Johnson-Neyman plot
sim_slopes(lag_ins_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM, jnplot = TRUE)

#save Johnson-Neyman plot
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_ins_age_tf_int_JNplot.png")

#create interaction plot (make it pretty later)
#basic plot
Ins_age_tf_int_plot <- interact_plot(lag_ins_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM)
Ins_age_tf_int_plot

#linearity check
Ins_age_tf_int_plot_lin <- interact_plot(lag_ins_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM, modx.values="terciles", plot.points=TRUE, linearity.check=TRUE)
Ins_age_tf_int_plot_lin
#seems to be some non-linear effects - try adding quadratic term


```
##Test quadratic effects of age with tissue fraction covariate in LAG models
Interaction plots showed some nonlinear effects, so testing quadratic effects of age
```{r}

#create quadratic age variable
lag_dat$mri_age_y2<- (lag_dat$mri_age_y)^2

#NAA
#age + tissue fraction model
lag_naa_age_TisFrac <- lmer(NAA_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age_TisFrac)

#add quadratic age term with all interactions
lag_naa_age_TisFrac_quad <- lmer(NAA_conc_molal ~ mri_age_y * mri_age_y2 * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_naa_age_TisFrac_quad) #no sig. effects

anova(lag_naa_age_TisFrac, lag_naa_age_TisFrac_quad) #not sig.

#Cre
#age + tissue fraction model
lag_cre_age_TisFrac<- lmer(Cre_PCr_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age_TisFrac)

#add quadratic age term with all interactions
lag_cre_age_TisFrac_quad<- lmer(Cre_PCr_conc_molal ~ mri_age_y * mri_age_y2 * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cre_age_TisFrac_quad)

#compare models
anova(lag_cre_age_TisFrac, lag_cre_age_TisFrac_quad) #not sig.

#Cho

#age x tissue fraction interaction model
lag_cho_age_TisFrac_int<- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age_TisFrac_int) #interaction is sig. (p=.046)

#add quadratic age term with all interactions
lag_cho_age_TisFrac_int_quad <- lmer(Cho_GPC_PCh_conc_molal ~ mri_age_y * mri_age_y2 * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_cho_age_TisFrac_int_quad) #no sig. effects

#compare model fit
anova(lag_cho_age_TisFrac_int, lag_cho_age_TisFrac_int_quad) #not sig., stick with age x tissue fraction model

#Ins

#age x tissue fraction interaction model
lag_ins_age_TisFrac_int<- lmer(Ins_conc_molal ~ mri_age_y * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age_TisFrac_int) #tissue fraction, age & interaction all sig.

#add quadratic age term with all interactions
lag_ins_age_TisFrac_int_quad<- lmer(Ins_conc_molal ~ mri_age_y * mri_age_y2 * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_ins_age_TisFrac_int_quad) #no sig. effects

anova(lag_ins_age_TisFrac_int, lag_ins_age_TisFrac_int_quad) #not sig.

#Glx
#age + tissue fraction model
lag_glx_age_TisFrac<- lmer(Glu_Gln_conc_molal ~ mri_age_y + TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age_TisFrac)

#add quadratic age term with all interactions
lag_glx_age_TisFrac_quad<- lmer(Glu_Gln_conc_molal ~ mri_age_y * mri_age_y2 * TissueFraction_GMWM + (1 | subj_id), data=lag_dat, REML=FALSE)
summary(lag_glx_age_TisFrac_quad) #no sig. effects
#warning: boundary fit is singular

#compare models
anova(lag_glx_age_TisFrac, lag_glx_age_TisFrac_quad)  #significant 



```


##Print model summaries 
Print formatted model summaries and tables using report and sjplot package
```{r}
library(report)
#library(sjplot) #try for pretty summary tables

summary(lag_naa_age_TisFrac)
report(lag_naa_age_TisFrac)
report_table(lag_naa_age_TisFrac)


summary(lag_cre_age_TisFrac)
report(lag_cre_age_TisFrac)

summary(lag_cho_age_TisFrac_int)
report(lag_cho_age_TisFrac_int)


```
##Plot results of models including tissue fraction
Estimate metabolite residuals using remef
Include individual slopes (estimated via linear models by subject) using geom_smooth()
Include overall fit line using geom_abline
```{r}
#LAG - NAA

#pull fixed effects for overall model
fixef.lag_naa_TisFrac<-fixef(lag_naa_age_TisFrac)

#subset dataframe with only subjects included in NAA analysis
lag_dat_plot<- lag_dat %>% 
  filter(! is.na(NAA_conc_molal))

#calculate partial effect of age on tNAA (remove effect of tissue fraction)
lag_naa_partial <- remef(lag_naa_age_TisFrac, fix = "TissueFraction_GMWM") #does NOT remove random effects variance

#create plot to show NAA ~ age while controlling for Tissue Fraction
plot_lag_naa_age_tf<-ggplot(lag_dat_plot, aes(x=mri_age_y, y=lag_naa_partial, na.exclude = TRUE, color=as.factor(female))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, aes(x=mri_age_y, y=lag_naa_partial, group=as.factor(subj_id)), size = .2, show.legend = FALSE) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "tNAA | Tissue Fraction") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_naa_TisFrac[[1]], slope = fixef.lag_naa_TisFrac[[2]])) + 
  theme_classic()
plot_lag_naa_age_tf
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_naa_age_partialef.png")


#LAG - Cre

#pull fixed effects for overall model
fixef.lag_cre_age_TisFrac<-fixef(lag_cre_age_TisFrac)

#subset dataframe with only subjects included in NAA analysis
lag_dat_plot_Cre<- lag_dat %>% 
  filter(! is.na(Cre_PCr_conc_molal))

#calculate partial effect of age on Cre (remove effect of tissue fraction)
lag_cre_partial <- remef(lag_cre_age_TisFrac, fix = "TissueFraction_GMWM") #does NOT remove random effects variance

plot_lag_cre_age_tf<-ggplot(lag_dat_plot_Cre, aes(x=mri_age_y, y=lag_cre_partial, na.exclude = TRUE, color=as.factor(female))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, aes(x=mri_age_y, y=lag_cre_partial, group=as.factor(subj_id)), size = .2, show.legend = FALSE) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "tCre | Tissue Fraction") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_cre_age_TisFrac[[1]], slope = fixef.lag_cre_age_TisFrac[[2]])) + 
  theme_classic()
plot_lag_cre_age_tf
  ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_cre_age_partialef.png")


#LAG - Cho
Cho_age_tf_int_plot <- interact_plot(lag_cho_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM, data=lag_dat, plot.points=TRUE, legend.main="Tissue Fraction GM/(GM+WM)", modx.labels=c(".535 (-1 SD)", ".642 (Mean)", ".750 (+1 SD)")) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "tCho (molal)") + 
  scale_color_viridis(option="mako", name="Tissue Fraction GM/(GM+WM)", begin= .32, end=1) +
  theme_classic()
Cho_age_tf_int_plot
  ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_cho_age_tf_int.png")


#LAG - Ins
#Main effect of age 
 #pull fixed effects for overall model
fixef.lag_ins_age_TisFrac_int<-fixef(lag_ins_age_TisFrac_int)

#subset dataframe with only subjects included in NAA analysis
lag_dat_plot_Ins<- lag_dat %>% 
  filter(! is.na(Ins_conc_molal))

#calculate partial effect of age on Ins (remove effect of tissue fraction)
lag_ins_partial <- remef(lag_ins_age_TisFrac_int, fix = c("TissueFraction_GMWM","mri_age_y:TissueFraction_GMWM")) #does NOT remove random effects variance

#plot main effect of age controlling for tissue fraction and age x tissue fraction interaction
plot_lag_ins_age_tf<-ggplot(lag_dat_plot_Ins, aes(x=mri_age_y, y=lag_ins_partial, na.exclude = TRUE, color=as.factor(female))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, aes(x=mri_age_y, y=lag_ins_partial, group=as.factor(subj_id)), size = .2, show.legend = FALSE) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "Ins | Tissue Fraction") +
  scale_color_viridis(name = "Sex", labels = c("Male", "Female"), discrete = TRUE, option = "H", begin = .2, end = .8) +
  geom_abline(aes(intercept = fixef.lag_ins_age_TisFrac_int[[1]], slope = fixef.lag_ins_age_TisFrac_int[[2]])) + 
  theme_classic()
plot_lag_ins_age_tf
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_ins_age_tf_maineffect.png")

  
#age x tissue fraction interaction plot
Ins_age_tf_int_plot <- interact_plot(lag_ins_age_TisFrac_int, pred = mri_age_y, modx = TissueFraction_GMWM, data=lag_dat, plot.points=TRUE, legend.main="Tissue Fraction GM/(GM+WM)", modx.labels=c(".535 (-1 SD)", ".642 (Mean)", ".750 (+1 SD)")) +
  labs(title = "Left Temporo-Parietal", x = "Age (years)", y = "tCre (molal)") +
  scale_color_viridis(option="mako", name="Tissue Fraction GM/(GM+WM)", begin= .32, end=1) +
  theme_classic()
Ins_age_tf_int_plot #only the slope for the lowest range of tissue fraction values (-1 SD) is significant 
ggsave("/Users/meaghan/OneDrive - University of Calgary/Preschool_data/MRS/results/lag_ins_age_tf_int.png")


```

